Analyzing X-ray images in Python with scikit-image

### 2016-08-21 skimage_tomo.tex
More edits

The Jupyter notebook \citep{Kluyver2016} is a web application that grew
from the IPython project. Jupyter notebooks provide a development
environment within a web browser, where live code can be enriched
by explanatory text, equations and visualizations (FIGURE).
As of July 2016, more than 500,000 Jupyter notebooks were posted on GitHub,
demonstrating their wide adoption by the community as workflow-sharing
tools (http://archive.ipython.org/media/SciPy2016JupyterLab.pdf).

@@

Visualizing images is an important component of the image processing
workflow, in order to verify the final result and to adjust the
parameters of intermediate processing operations.
\texttt{matplotlib} \citep{Hunter2007} is the most popular 2D plotting
library of the Python ecosystem. It can be used to visualize 2D data such
as color or grayscale images, and 1D data such as contour lines, outlines
of segmented regions, histograms of gray levels, etc. Although
\texttt{matplotlib} has simple 3D plotting capabilities, we
recommend using the \texttt{mayavi} module \citep{Ramachandran2011}
for applications requiring advanced 3D visualization, such as tomography. 
\texttt{mayavi} is based on the VTK toolkit. It exposes a simple API for
visualizing data passed as \texttt{numpy} arrays. For more advanced
visualizations, a large majority of VTK capabilities can be accessed
thanks to a pipeline API.
\texttt{mayavi} offers a good trade-off between simplicity of use for
common operations, and more advanced possibilities, including responsive
visualizations.

@@

\texttt{scikit-image} is only one Python module that can be used for data
processing, among many others. A very popular module is
\texttt{scikit-learn} \citep{Pedregosa2011}, a Python module for machine
learning using \texttt{NumPy} arrays. If local features of an image
(such as local statistics of gray levels, or geometric points of
interest), or features of segmented objects (e.g. geometrical and
intensity characteristics of segmented particles) are extracted with
functions from \texttt{skimage.feature} (see Fig.~\ref{fig:ecosystem}), it
is then possible to use a \emph{classification} algorithm of
\texttt{scikit-learn}, either for labeling pixels (a segmentation task)
or to classify whole images or objects that have already been segmented.
The interoperability between these packages is ensured by the common use
of \texttt{NumPy} arrays, so that the tasks described above are
implemented with a few lines of code.

The modularity of the Scientific Python ecosystem may be
confusing at first sight. Nevertheless, core modules of this ecosystem
are almost perfectly compatible, thanks to the shared use of NumPy arrays
and common development practices (although they are developed in parallel
by different teams). Several integrated distributions, such as Anaconda
or Canopy, bundle together the most common packages, including
\texttt{scikit-image}.

@@

\begin{figure*}
    \centerline{\includegraphics[width=0.99\textwidth]{tomo_gallery}}
    \caption{\csentence{Typical image processing operations with
	\texttt{scikit-image}.} From left to right and top to bottom.
	Data are synthetic, unless stated otherwise.
	\textbf{Filtering -} Top: non-local means denoising of an image
	with a fine-grained texture, acquired by \emph{in situ}
	synchrotron microtomography during
	glass melting \citep{Gouillart2012}. Bottom: total-variation
	denoising of an image with two phases, corresponding to
	phase-separating silicate melts observed by \emph{in situ
	tomography} \citep{Bouttes2015}.
	\textbf{Feature extraction -} Top: Hubble deep field (NASA,
	public domain), blob detection using the
	Laplacian of Gaussian method. Bottom: ridge detection using the
	leading value of the Hessian matrix. 
	\textbf{Segmentation - } Top: super-pixel segmentation
	of a CT slice of the human head \citep{tomo_wikipedia}, using
	Felzenszwalb's algorithm \citep{Felzenszwalb2004}. Bottom: random
	walker segmentation (right) of noisy image (top-left corner), using
	histogram-determined markers (bottom-left corner).
	\textbf{Measures -} Top: visualization of local diameter
	(color-coded on the skeleton curve) of an
	interconnected phase (represented in violet).  Bottom: particles color-coded according
	to their extent.
\label{fig:tomo_gallery}}
\end{figure*}

@@

Segmentation of regions of interest can be achieved using different
strategies, depending on the characteristics of the image. Images with a
clear contrast between regions can be segmented automatically thanks to
several thresholding algorithms, including an adaptive local thresholding
algorithm aimed at images with contrast variations. Super-pixel
algorithms \citep{Felzenszwalb2004, Achanta2012} (see
Fig.~\ref{fig:tomo_gallery}) create an over-segmentation of images in
super-pixels, by grouping pixels that are close together both in color-
and direct space. Region-growing algorithms, such as the morphological
watershed or the random walker \citep{Grady2006}, propagate the labels
of user-defined markers through the image (see
Fig.~\ref{fig:tomo_gallery}). The active contour
algorithm \citep{Kass1988} fits snake contours to features of the
image, such as edges or high-brightness regions.

@@

On a different note, prototyping the image processing workflow requires several trial and error
executions of scripts containing different operations. The \texttt{joblib}
module provides a caching mechanism that avoids the repetition of function
calls, if their arguments have not changed:
\begin{lstlisting}
>>> from skimage import filters, data
>>> from joblib import Memory
>>> mem = Memory(cachedir='/tmp/joblib')
>>> median = mem.cache(filters.median)
>>> im = data.camera()
>>> filtered_im = median(im, np.ones((3, 3)))
__________________________
[Memory] Calling skimage.filters.rank.generic.median...
median(array([[156, ..., 152],
       ..., 
       [121, ..., 111]], dtype=uint8), array([[ 1.,  1.,  1.],
       [ 1.,  1.,  1.],
       [ 1.,  1.,  1.]]))
__________________________median - 0.0s, 0.0min
>>> filtered_again = median(im, np.ones((3, 3)))
>>> # the above call did not trigger an evaluation
\end{lstlisting}

Finally, let us note that the biggest performance improvements often come from
the improvement of algorithms (as opposed to computing architectures only). For
example, non-local means denoising \citep{Buades2005} is a costly operation,
since it requires several nested loops, on all pixels and on neighboring
patches to be compared with the pixel-centered patch. Thanks to the
implementation of a more recent algorithm \citep{Darbon2008} that modifies the
internal organization of loops, it was possible to gain a factor of 5 to 10.
The large size of the \texttt{scikit-image} community ensures that possible
algorithmic improvements are discussed regularly.

@@

The quality of software documentation is an important component to the
success of a package aimed at scientists.
\texttt{scikit-image} users have access to different kinds of
documentation. Functions of the API are all
documented using the same standard \citep{Pawlik2015} as other major
Scientific Python packages. The standard include a description of all
input and output variables and their data type, together with
explanations on what the function does and how to use it. Function
documentation is accessed either online, or within the development
environment (IPython, Spyder, Jupyter Notebook...).

\begin{figure*}
    \centerline{\includegraphics[width=0.99\textwidth]{figure_gallery}}
\caption{\csentence{Gallery of examples of \texttt{scikit-image}.}
 The gallery of examples consists of an array of thumbnails (left part), which link to example webpages, each centerered on a specific image processing task. Each webpage includes Python code generating a figure, the figure itself, as well as a short tutorial explaining the image processing operations and the code. \label{fig:gallery}}
\end{figure*}

In addition, a popular documentation tool is the graphical gallery of
examples (\nolinkurl{http://scikit-image.org/docs/dev/auto_examples/}), a part of which is displayed in Fig.~\ref{fig:gallery}.
The principle of the gallery is to accelerate the learning curve of image
processing, by showcasing graphical examples of image processing
operations. Such examples are organized as an array of thumbnails with a
short title (see Fig.~\ref{fig:gallery} left). These thumbnails link to
the webpage of the corresponding example, which features a mini-tutorial
on the image processing method, the code needed to run the example and
the figure generated by the example. Since the graphical gallery is an
efficient way to inform users about the features of
\texttt{scikit-image}, every new feature integrated in the package must
include an example for the gallery.
Longer tutorials and a more narrative documentation is available as well
in the online User Guide of \texttt{scikit-image}. The User Guide explains in
particular "big picture" aspects of \texttt{scikit-image}, such as its
use of \texttt{NumPy} arrays as images, or how the package interacts with
other parts of the scientific Python ecosystem.

@@

Estimating the number of active users of an open-source package is a
difficult task. Statistics of downloads, for example, largely
overestimate the number of active users, all the more if the package is
bundled with others in a software distribution, like Anaconda or
Canopy. A view closer to reality can be obtained by analyzing the
statistics of visits of the online help, available on the project
website.  As of the first half of 2016, 20000 unique visitors visit the
website of scikit-image every month \url{http://scikit-image.org/}. These
users come from 80 countries, located almost equally in Europe, America
and Asia (Oceania and Africa representing a much smaller fraction of
visitors).

@@

\texttt{scikit-image} is developed by a diverse team of volunteers, with
more than 170 developers that have contributed to the package. The large
number of developers and users is a key element of the sustainability of
\texttt{scikit-image}. The
development process takes place on the GitHub platform
\url{https://github.com/scikit-image/scikit-image}, where users and
developers propose and discuss new contributions, or report bugs or ideas
for improvements.
A release cycle of one or two releases every year
ensures that new features are proposed to users on a regular basis.

@@

\texttt{scikit-image} offers a wide variety of image processing
algorithms, using a simple interface natively compatible with 2D and 3D
images. It is well integrated into the Scientific Python ecosystem, so
that it interfaces well with visualization libraries or other data
processing packages. \texttt{scikit-image} has seen a tremendous growth
since its creation in 2010, both in terms of users and proposed features.
In addition to the growing number of scientific teams that use
\texttt{scikit-image} for processing X-ray-modalities images,
domain-specific tools are now using \texttt{scikit-image} as a dependency
to build upon. Examples include \texttt{tomopy} \citep{Gursoy2014} for
tomographic reconstruction, or \texttt{DIOPTAS} \citep{Prescher2015} for
the reduction and exploration of X-ray diffraction data. It is likely
that more application-specific softwares will benefit from depending on
\texttt{scikit-image} in the future, since \texttt{scikit-image} strives
to be domain-agnostic and to keep a stable interface. On the end-user
side, future work includes a better integration of parallel processing
capabilities, completion of full 3-D compatibility, an enriched narrative
documentation, and the extension of the number of algorithms.

>>>>

The Jupyter notebook \citep{Kluyver2016} is a web application that grew
from the IPython project. Jupyter notebooks provide a development
environment within a web browser, where live code can be enriched
by explanatory text, equations and visualizations (FIGURE).
As of July 2016, more than 500,000 Jupyter notebooks were posted on GitHub,
demonstrating their wide adoption by the community as workflow-sharing
tools (\url{http://archive.ipython.org/media/SciPy2016JupyterLab.pdf}).

@@

Visualizing images is an important component of the image processing
workflow, in order to verify the final result and to adjust the
parameters of intermediate processing operations.
\texttt{matplotlib} \citep{Hunter2007} is the most popular 2D plotting
library of the Python ecosystem. It can be used to visualize 2D data such
as color or grayscale images, and 1D data such as contour lines, outlines
of segmented regions, histograms of gray levels, etc. Although
\texttt{matplotlib} has simple 3D plotting capabilities, we
recommend using the \texttt{mayavi} module \citep{Ramachandran2011}
for applications requiring advanced 3D visualization, such as tomography. 
\texttt{mayavi} is based on the VTK toolkit. It exposes a simple API for
visualizing data passed as \texttt{numpy} arrays. For more advanced
visualizations, a large majority of VTK capabilities can be accessed
through its pipeline API.
\texttt{mayavi} offers a good trade-off between simplicity of use for
common operations, and more advanced possibilities, including responsive
visualizations.

@@

\texttt{scikit-image} is only one Python module that can be used for data
processing, among many others. A very popular module is
\texttt{scikit-learn} \citep{Pedregosa2011}, a Python module for machine
learning using \texttt{NumPy} arrays. If local features of an image
(such as local statistics of gray levels, or geometric points of
interest), or features of segmented objects (e.g. geometrical and
intensity characteristics of segmented particles) are extracted with
functions from \texttt{skimage.feature} (see Fig.~\ref{fig:ecosystem}), it
is then possible to use a \emph{classification} algorithm of
\texttt{scikit-learn}, either for labeling pixels (a segmentation task)
or to classify whole images or objects that have already been segmented.
The near-universal use of \texttt{NumPy} arrays ensures the interoperability
between these packages, so that just a few lines of code are sufficient
to create these sophisticated workflows.

The modularity of the Scientific Python ecosystem may be
confusing at first sight, but the core modules of this ecosystem
are almost perfectly compatible, thanks to the shared use of NumPy arrays
and common development practices (although they are developed in parallel
by different teams). Several ``distributions'', such as Anaconda
or Canopy, bundle together the most popular libraries, including
\texttt{scikit-image}.

@@

\begin{figure*}
    \centerline{\includegraphics[width=0.99\textwidth]{tomo_gallery}}
    \caption{\csentence{Typical image processing operations with
	\texttt{scikit-image}.} From left to right and top to bottom.
	Data are synthetic, unless stated otherwise.
	\textbf{Filtering -} Top: non-local means denoising of an image
	with a fine-grained texture, acquired by \emph{in situ}
	synchrotron microtomography during
	glass melting \citep{Gouillart2012}. Bottom: total-variation
	denoising of an image with two phases, corresponding to
	phase-separating silicate melts observed by \emph{in situ
	tomography} \citep{Bouttes2015}.
	\textbf{Feature extraction -} Top: Hubble deep field (NASA,
	public domain), blob detection using the
	Laplacian of Gaussian method. Bottom: ridge detection using the
	leading eigenvalue of the Hessian matrix.
	\textbf{Segmentation - } Top: super-pixel segmentation
	of a CT slice of the human head \citep{tomo_wikipedia}, using
	Felzenszwalb's algorithm \citep{Felzenszwalb2004}. Bottom: random
	walker segmentation (right) of noisy image (top-left corner), using
	histogram-determined markers (bottom-left corner).
	\textbf{Measures -} Top: visualization of local diameter
	(color-coded on the skeleton curve) of an
	interconnected phase (represented in violet).  Bottom: particles color-coded according
	to their extent.
\label{fig:tomo_gallery}}
\end{figure*}

@@

Segmentation of regions of interest can be achieved using different
strategies, depending on the characteristics of the image. Images with a
clear contrast between regions can be segmented automatically thanks to
several thresholding algorithms, including an adaptive local thresholding
algorithm aimed at images with contrast variations. Super-pixel
algorithms \citep{Felzenszwalb2004, Achanta2012} create an over-segmentation
of images in
super-pixels, by grouping pixels that are close together both in color-
and direct space (see Fig.~\ref{fig:tomo_gallery}). Region-growing
algorithms, such as the morphological
watershed or the random walker \citep{Grady2006}, propagate the labels
of user-defined markers through the image (see
Fig.~\ref{fig:tomo_gallery}). The active contour
algorithm \citep{Kass1988} fits snake contours to features of the
image, such as edges or high-brightness regions.

@@

\em{Caching} provides another tool to speed up data analysis. Prototyping an
image processing workflow often requires several trial and error
executions of scripts containing different operations. \texttt{joblib}
provides a caching mechanism that avoids the repetition of function
calls, if their arguments have not changed:
\begin{lstlisting}
>>> from skimage import filters, data
>>> from joblib import Memory
>>> mem = Memory(cachedir='/tmp/joblib')
>>> median = mem.cache(filters.median)
>>> im = data.camera()
>>> filtered_im = median(im, np.ones((3, 3)))
__________________________
[Memory] Calling skimage.filters.rank.generic.median...
median(array([[156, ..., 152],
       ..., 
       [121, ..., 111]], dtype=uint8), array([[ 1.,  1.,  1.],
       [ 1.,  1.,  1.],
       [ 1.,  1.,  1.]]))
__________________________median - 0.0s, 0.0min
>>> filtered_again = median(im, np.ones((3, 3)))
>>> # the above call did not trigger an evaluation
\end{lstlisting}

Finally, we note that the biggest performance improvements often come from
the improvement of algorithms (as opposed to computing architectures only). For
example, non-local means denoising \citep{Buades2005} is a costly operation,
since it requires several nested loops, on all pixels and on neighboring
patches to be compared with the pixel-centered patch. Thanks to the
implementation of a more recent algorithm \citep{Darbon2008} that modifies the
internal organization of loops, it was possible to gain a 5- to 10-fold speedup.
The large size of the \texttt{scikit-image} community ensures that possible
algorithmic improvements are discussed regularly.

@@

The quality of the documentation is an important component to the
success of a software package.
\texttt{scikit-image} users have access to different kinds of
documentation. All functions are
documented using the NumPy documentation standard \citep{Pawlik2015},
which is universal across all major
Scientific Python packages. The standard includes a description of all
input and output variables and their data type, together with
explanations on what the function does and how to use it. Function
documentation is accessible online or within the development
environment itself (IPython, Spyder, Jupyter Notebook...).

\begin{figure*}
    \centerline{\includegraphics[width=0.99\textwidth]{figure_gallery}}
\caption{\csentence{Gallery of examples of \texttt{scikit-image}.}
 The gallery of examples consists of an array of thumbnails (left), which link to example webpages, each centerered on a specific image processing task. Each webpage includes Python code generating a figure, the figure itself, and a short tutorial explaining the image processing operations and the code. \label{fig:gallery}}
\end{figure*}

In addition, a popular documentation tool is the graphical gallery of
examples (\nolinkurl{http://scikit-image.org/docs/dev/auto_examples/}),
an example of which appears in Fig.~\ref{fig:gallery}.
The principle of the gallery is to accelerate the learning curve of image
processing, by showcasing graphical examples of common image processing
operations. Such examples are organized as an array of thumbnails with a
short title (see Fig.~\ref{fig:gallery} left). These thumbnails link to
the webpage of the corresponding example, which features a mini-tutorial
on the image processing method, the code needed to run the example and
the figure generated by the example. Since the graphical gallery is an
efficient way to inform users about the features of
\texttt{scikit-image}, every new feature integrated in the package must
include an example for the gallery.
Longer tutorials and a more narrative documentation is available as well
in the online User Guide, which explains
"big picture" aspects of \texttt{scikit-image}, such as its
use of \texttt{NumPy} arrays as images, or how the package interacts with
other parts of the scientific Python ecosystem.

@@

Estimating the number of active users of an open-source package is a
difficult task. Statistics of downloads, for example, largely
overestimate the number of active users, all the more if the package is
bundled with others in a software distribution, such as Anaconda or
Canopy. A view closer to reality can be obtained by analyzing the
statistics of visits of the online help, available on the project
website.  As of the first half of 2016, 20000 unique visitors visit the
website of scikit-image every month \url{http://scikit-image.org/}. These
users come from 80 countries, located almost equally in Europe, America
and Asia (with Oceania and Africa representing a much smaller fraction of
visitors).

@@

\texttt{scikit-image} is developed by a diverse team of volunteers.
More than 170 individuals have contributed to the package. The large
number of developers and users is a key element of the sustainability of
\texttt{scikit-image}. The
development process takes place on GitHub
\url{https://github.com/scikit-image/scikit-image}, where users and
developers propose and discuss new contributions, or report bugs or ideas
for improvements.
A release cycle of one or two releases every year
ensures that new features are proposed to users on a regular basis.

@@

\texttt{scikit-image} offers a wide variety of image processing
algorithms, using a simple interface natively compatible with 2D and 3D
images. It is well integrated into the Scientific Python ecosystem, so
that it interfaces well with visualization libraries or other data
processing packages. \texttt{scikit-image} has seen a tremendous growth
since its creation in 2010, both in terms of users and proposed features.
In addition to the growing number of scientific teams that use
\texttt{scikit-image} for processing X-ray-modalities images,
domain-specific tools are now using \texttt{scikit-image} as a dependency
to build upon. Examples include \texttt{tomopy} \citep{Gursoy2014} for
tomographic reconstruction, or \texttt{DIOPTAS} \citep{Prescher2015} for
the reduction and exploration of X-ray diffraction data. It is likely
that more application-specific software will benefit from depending on
\texttt{scikit-image} in the future, since \texttt{scikit-image} strives
to be domain-agnostic and to keep a stable interface. On the end-user
side, future work includes a better integration of parallel processing
capabilities, completion of full 3-D compatibility, an enriched narrative
documentation, and the extension of the number of algorithms.
